---
title: "Veteran's Life History and Psychological Health"
subtitle: "A dataset from San Diego VA hospital"
author: "Shuai Shao, Shuyuan Shi, Xinyue Li"
date: "11/14/2019"
output: html_document
---

In this group project, we will use github to record and track every team member's work. You can find our project at this [github repository](https://github.com/shaoshuai95/PSYC201A-project.git).

For this project, we are aiming to explore the veteran’s psychological health patterns in categories of depression, anxiety, sleep patterns, social connections, drug abuse, and social support. 

In order to obtain data of veterans, we have got permission to access the dataset that records veterans’ treatment records from the VA hospital. Currently, we have demographic information (age, gender, ethnicity, social relationship status), alcohol usage, anxiety disorder scale, depression scale, emotional support, sleep patterns, BMI, and military service deployment (type, branch, number).

After conducting some preliminary analyses, we plan to gather more variables such as veteran's social support, traumatic experience (such as childhood trauma and trauma during military service), substance abuse other than alcohol so that we can have a more overall analysis of the psychological health for veterans in the VA hospital. We can get necessary dataset by November 20th. 

We plan to analyze:  
1) The relationship between veterans’ emotional support status and their anxiety/depression level (linear regression, correlation).  
2) The relationship between traumatic experiences, their substance abuse, and their anxiety levels (linear regression, multiple regression, correlation).  
3) The possible difference in types and branches of military service deployment and the veteran’s traumatic experiences (ANOVA, chi-squared test).  
4) The possible differences of all mentioned variables in gender, age, and ethnicity (linear regression, chi-squared test, t-test).  
5) We will also explore other possible relationships of the data.  

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
```

```{r import data}
data <- readxl::read_xlsx("PSYC201a.xlsx")
```

```{r clean the data}
# select varables of interest
data <- data %>% 
  select(-2,-12,-13)

# rename the columns
data <- rename(data,"gender"="demo_gender",
                    "Hispanic"="demo_ethnic",
                    "Caucasian"="demo_racewhite",
                    "African"="demo_race_black",
                    "Native"="demo_race_amind",
                    "Pacific_Island"="demo_race_pacisl",
                    "Asian"="demo_race_asian",
                    "Unclassified"="demo_race_decline",
                    "age"="Calculate999age",
                    "relationship"="demo_relationship",
                    "ServiceType"="serv_type_0",
                    "ServiceBranch"="serv_branch_0",
                    "ServiceDeployCount"="serv_deploy_count",
                    "anxiety"="gad7_score",
                    "sleep quality"="sleep_score",
                    "depression"="dep_score_phq9",
                    "bmi"="demo_BMI",
                    "emotion support"="es_score_promis",
                    "income"="demo_income_group",
                    "education"="demo_education",
                    "work status"="demo_workstatus",
                    "employed hours"="demo_hours",
                    "alone"="demo_livewith_alone",
                    "parent"="demo_livewith_parent",
                    "friend"="demo_livewith_friend",
                    "child"="demo_livewith_child",
                    "other"="demo_livewith_other",
                    "none support"="demo_emo_none",
                    "spouse support"="demo_emo_spouse",
                    "therapist support"="demo_emo_therapist",
                    "spiritual support"="demo_emo_spiritual",
                    "children support"="demo_emo_children",
                    "other support"="demo_emo_other",
                    "relationship hurt"="demo_rel_hurt",
                    "alcohol"="alc_score_audit") 
```

```{r mutate the variable format}
data <- data %>%
  mutate (Hispanic=as.character(Hispanic),
          Caucasian=as.character(Caucasian),
          African=as.character(African),
          Native=as.character(Native),
          Pacific_Island=as.character(Pacific_Island),
          Asian=as.character(Asian),
          Unclassified=as.character(Unclassified))
```

```{r dealing with missing data}
# replace all "999"s with "NA"
data[data==999]=NA

data <- data %>%
  filter(!is.na(gender)) %>%
  filter(!is.na(Caucasian)) %>%
  filter(!is.na(Hispanic)) %>%
  filter(!is.na(African)) %>%
  filter(!is.na(Native)) %>%
  filter(!is.na(Pacific_Island)) %>%
  filter(!is.na(Asian)) %>%
  filter(!is.na(Unclassified)) 

```

```{r recode ethnics}
data <- unite(data, Ethnics, Hispanic, Caucasian, African, Native, Pacific_Island, Asian, Unclassified, sep="", remove=FALSE)

data$Ethnics[data$Ethnics=="1000000"]="Hispanic"
data$Ethnics[data$Ethnics=="0100000"]="Caucasian"
data$Ethnics[data$Ethnics=="0010000"]="African"
data$Ethnics[data$Ethnics=="0001000"]="Native"
data$Ethnics[data$Ethnics=="0000100"]="Pacific Island"
data$Ethnics[data$Ethnics=="0000010"]="Asian"
data$Ethnics[data$Ethnics=="0000001"]="Unclassified"

data$Ethnics[!(data$Ethnics %in% c("Hispanic","Caucasian","African","Native","Pacific Island","Asian","Unclassified"))]="Other"
```

```{r recode living status}
data <- unite(data, support, alone, parent, friend, child, other, sep="", remove=FALSE)
data$support[data$support=="10000"]="alone"
data$support[data$support=="01000"]="parent"
data$support[data$support=="00100"]="friend"
data$support[data$support=="00010"]="child"

data$support[!(data$support %in% c("alone","parent","friend","child"))]="other"
```

```{r filter out unreasonable cases}
# filter out unreasonable cases
data <- data %>% 
  filter(age<100) %>%
  filter(age>18)
```

```{r rename values}
#recode gender
data$gender[data$gender==2]="female"
data$gender[data$gender==1]="male"

#recode relationships
data$relationship[data$relationship==1]="single"
data$relationship[data$relationship==2]="married"
data$relationship[data$relationship==3]="seperated"
data$relationship[data$relationship==4]="divorced"
data$relationship[data$relationship==5]="cohabitating"
data$relationship[data$relationship==6]="civil union"
data$relationship[data$relationship==7]="remarried"
data$relationship[data$relationship==8]="widowed"

#recode service type
data$ServiceType = as.factor(data$ServiceType)
data$ServiceType[data$ServiceType == "1"] = "Active Duty"
data$ServiceType[data$ServiceType == "2"] = "Reserve"
data$ServiceType[data$ServiceType == "3"] = "Guard"

#recode service branch
data$ServiceBranch = as.factor(data$ServiceBranch)
data$ServiceBranch[data$ServiceBranch == "1"] = "Army"
data$ServiceBranch[data$ServiceBranch == "2"] = "Air Force"
data$ServiceBranch[data$ServiceBranch == "3"] = "Coast Guard"
data$ServiceBranch[data$ServiceBranch == "4"] = "Marines"
data$ServiceBranch[data$ServiceBranch == "5"] = "National"
data$ServiceBranch[data$ServiceBranch == "6"] = "Navy"

#recode income
data$income = as.character(data$income)
data$income[data$income == "1"] = "less than 15k"
data$income[data$income == "2"] = "15k~30k"
data$income[data$income == "3"] = "30k~45k"
data$income[data$income == "4"] = "45k~60k"
data$income[data$income == "5"] = "60k~75k"
data$income[data$income == "6"] = "75k~100k"
data$income[data$income == "7"] = "more than 100k"

#recode education
data$education = as.character(data$education)
data$education[data$education == "1"] = "some high school"
data$education[data$education == "2"] = "GED"
data$education[data$education == "3"] = "high school"
data$education[data$education == "4"] = "some college"
data$education[data$education == "5"] = "associates degree"
data$education[data$education == "6"] = "college degree"
data$education[data$education == "7"] = "master's degree"
data$education[data$education == "8"] = "doctoral degree"

#recode work status
data$`work status` = as.character(data$`work status`)
data$`work status`[data$`work status` == "1"] = "full time"
data$`work status`[data$`work status` == "2"] = "part time"
data$`work status`[data$`work status` == "3"] = "seasonally"
data$`work status`[data$`work status` == "4"] = "day labor"
data$`work status`[data$`work status` == "5"] = "unemployed"
```

## Demographical Variables

```{r demographical variables}
data %>%
  ggplot() +
  geom_bar(aes(x=Ethnics,fill=gender),stat="count",position = "dodge")+
  theme_bw() +
  labs(title = "Distribution of Ethnicity by Gender",
       x = "Ethnicity",
       y = "Count")
```


## Preliminary analyses
```{r sleeping quality~relationship+gender}
data %>%
  ggplot()+
  geom_boxplot(aes(x=relationship,y = `sleep quality`,color = gender)) +
  labs(title = "Relationship Between Relationship Status And Sleep",
       x = "Relationship Status",
       y = "Sleep Quality") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))

model_rel_sleep=lm(`sleep quality`~relationship,data)
anova(model_rel_sleep)
```

```{r anxiety~relationship+gender}
data %>%
  ggplot() + 
  geom_boxplot(aes(x=relationship,y = anxiety, color = gender)) +
  labs(title = "Relationship Between Relationship Status And Anxiety",
       x = "Relationship Status",
       y = "Anxiety Level") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))

model_rel_anx=lm(data=data,formula=anxiety~relationship+gender)
anova(model_rel_anx)
```

```{r depression~relationship+gender}
data %>%  
  ggplot() + 
  geom_boxplot(aes(x=Relationship,y = dep_score_phq9, color = gender)) +
  labs(title = "Relationship Between Relationship Status And Depression",
       x = "Relationship Status",
       y = "Depression") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))

model_rel_dep=lm(data=data,formula=gad7_score~Relationship+gender)
anova(model_rel_dep)
```

```{r alcohol abuse~relationship+gender}
data %>%  
  ggplot() + 
  geom_boxplot(aes(x=relationship,y = alcohol, color = gender)) +
  labs(title = "Relationship Between Relationship Status And Alcohol Abuse",
       x = "Relationship Status",
       y = "Alcohol Abuse") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))

model_rel_alc=lm(data=data,formula=alcohol~relationship+gender)
anova(model_rel_alc)
```

```{r sleep quality~emotional support}
data %>%
  ggplot(aes(x=es_score_promis,y = sleep_score, color = gender))+
  geom_jitter(alpha=0.4, size=0.5)+
  geom_smooth()+
  theme_bw()+
   labs(title = "Relationship Between Emotional Support And Sleep",
       x = "Emotional Support",
       y = "Sleep")

cor.test(data$es_score_promis,data$sleep_score)

model_es_gender_sleep=lm(data=data,formula=sleep_score~es_score_promis+gender)
anova(model_es_gender_sleep)
```

```{r anxiety~emotional support}
data %>%
  ggplot(aes(x=es_score_promis,y=gad7_score,color=gender))+
  geom_jitter(alpha=0.4, size=0.5)+
  geom_smooth()+
  theme_bw()+
  labs(title = "Relationship Between Emotional Support And Anxiety",
       x = "Emotional Support",
       y = "Anxiety")

cor.test(data$es_score_promis,data$gad7_score)

model_es_gender_anx=lm(data=data,formula=gad7_score~es_score_promis+gender)
anova(model_es_gender_anx)
```

## Future Directions
In the following stage, we will  
1) add more variables of interest   
2) explore more complex relationships such as interactions among explanatory variables  
3) explain the results in APA style.