---
title: "Veteran's Life History and Psychological Health"
subtitle: "A dataset from San Diego VA hospital"
author: "Shuai Shao, Shuyuan Shi, Xinyue Li"
date: "12/12/2019"
output: html_document
---

In this group project, we used github to record and track every team member's work. You can find our project at this [github repository](https://github.com/shaoshuai95/PSYC201A-project.git).    
    
We aim to explore the veterans' psychological health patterns in categories of depression, anxiety, sleep patterns, and alcohol usage. We have got permission to access the dataset that records veterans’ treatment records from San Diego VA hospital.    
    
We hypothesize that:   
*Higher social support, indicated by relationship status and emotional support score, is positively associated with lower mental health severity (i.e. Anxiety, Depression, Alcohol Use, and Insomnia).
*Higher SES, indicated by work status, education level and income level, is positively associated with lower mental health severity (i.e. Anxiety, Depression, Alcohol Use and Insomnia).

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
```

```{r import data}
#Import data
data <- readxl::read_xlsx("PSYC201a.xlsx")
```
```{r clean the dataset}
#Select varables of interest (Exclude meaningless variables)
data <- data %>% 
  select(-2,-12,-13)

#Rename the columns (variable names)
data <- rename(data,"gender"="demo_gender",
                    "Hispanic"="demo_ethnic",
                    "Caucasian"="demo_racewhite",
                    "African"="demo_race_black",
                    "Native"="demo_race_amind",
                    "Pacific_Island"="demo_race_pacisl",
                    "Asian"="demo_race_asian",
                    "Unclassified"="demo_race_decline",
                    "age"="Calculate999age",
                    "relationship"="demo_relationship",
                    "ServiceType"="serv_type_0",
                    "ServiceBranch"="serv_branch_0",
                    "ServiceDeployCount"="serv_deploy_count",
                    "anxiety"="gad7_score",
                    "sleep quality"="sleep_score",
                    "depression"="dep_score_phq9",
                    "bmi"="demo_BMI",
                    "emotion support"="es_score_promis",
                    "income"="demo_income_group",
                    "education"="demo_education",
                    "work status"="demo_workstatus",
                    "employed hours"="demo_hours",
                    "alone"="demo_livewith_alone",
                    "parent"="demo_livewith_parent",
                    "friend"="demo_livewith_friend",
                    "child"="demo_livewith_child",
                    "other"="demo_livewith_other",
                    "none support"="demo_emo_none",
                    "spouse support"="demo_emo_spouse",
                    "therapist support"="demo_emo_therapist",
                    "spiritual support"="demo_emo_spiritual",
                    "children support"="demo_emo_children",
                    "other support"="demo_emo_other",
                    "relationship hurt"="demo_rel_hurt",
                    "alcohol"="alc_score_audit") 

#Mute the demographic variables format
data <- data %>%
  mutate (Hispanic=as.character(Hispanic),
          Caucasian=as.character(Caucasian),
          African=as.character(African),
          Native=as.character(Native),
          Pacific_Island=as.character(Pacific_Island),
          Asian=as.character(Asian),
          Unclassified=as.character(Unclassified))

#Filter out unreasonable cases
data <- data %>% 
  filter(age<100) %>%
  filter(age>18)
```

```{r dealing with missing data}
#Replace all "999"s with "NA"
data[data==999]=NA

#Filter out participants that have missing values in critical demographic variables
data <- data %>%
  filter(!is.na(gender)) %>%
  filter(!is.na(Caucasian)) %>%
  filter(!is.na(Hispanic)) %>%
  filter(!is.na(African)) %>%
  filter(!is.na(Native)) %>%
  filter(!is.na(Pacific_Island)) %>%
  filter(!is.na(Asian)) %>%
  filter(!is.na(Unclassified)) 
```

```{r recode data values}
#Recode ethnics
data <- unite(data, Ethnics, Hispanic, Caucasian, African, Native, Pacific_Island, Asian, Unclassified, sep="", remove=FALSE)

data$Ethnics[data$Ethnics=="1000000"]="Hispanic"
data$Ethnics[data$Ethnics=="0100000"]="Caucasian"
data$Ethnics[data$Ethnics=="0010000"]="African"
data$Ethnics[data$Ethnics=="0001000"]="Native"
data$Ethnics[data$Ethnics=="0000100"]="Pacific Island"
data$Ethnics[data$Ethnics=="0000010"]="Asian"
data$Ethnics[data$Ethnics=="0000001"]="Unclassified"

data$Ethnics[!(data$Ethnics %in% c("Hispanic","Caucasian","African","Native","Pacific Island","Asian","Unclassified"))]="Other"

#recode living status
data <- unite(data, support, alone, parent, friend, child, other, sep="", remove=FALSE)
data$support[data$support=="10000"]="alone"
data$support[data$support=="01000"]="parent"
data$support[data$support=="00100"]="friend"
data$support[data$support=="00010"]="child"

data$support[!(data$support %in% c("alone","parent","friend","child"))]="other"

#recode gender
data$gender[data$gender==2]="female"
data$gender[data$gender==1]="male"

#recode relationships
data$relationship[data$relationship==1]="single"
data$relationship[data$relationship==2]="married"
data$relationship[data$relationship==3]="separated"
data$relationship[data$relationship==4]="divorced"
data$relationship[data$relationship==5]="cohabitating"
data$relationship[data$relationship==6]="civil union"
data$relationship[data$relationship==7]="remarried"
data$relationship[data$relationship==8]="widowed"

#recode service type
data$ServiceType[data$ServiceType == "1"] = "Active Duty"
data$ServiceType[data$ServiceType == "2"] = "Reserve"
data$ServiceType[data$ServiceType == "3"] = "Guard"

#recode service branch
data$ServiceBranch[data$ServiceBranch == "1"] = "Army"
data$ServiceBranch[data$ServiceBranch == "2"] = "Air Force"
data$ServiceBranch[data$ServiceBranch == "3"] = "Coast Guard"
data$ServiceBranch[data$ServiceBranch == "4"] = "Marines"
data$ServiceBranch[data$ServiceBranch == "5"] = "National"
data$ServiceBranch[data$ServiceBranch == "6"] = "Navy"

#recode income
data$income = as.character(data$income)
data$income[data$income == "1"] = "less than 15k"
data$income[data$income == "2"] = "15k~30k"
data$income[data$income == "3"] = "30k~45k"
data$income[data$income == "4"] = "45k~60k"
data$income[data$income == "5"] = "60k~75k"
data$income[data$income == "6"] = "75k~100k"
data$income[data$income == "7"] = "more than 100k"

#recode education
data$education = as.character(data$education)
data$education[data$education == "1"] = "some high school"
data$education[data$education == "2"] = "GED"
data$education[data$education == "3"] = "high school"
data$education[data$education == "4"] = "some college"
data$education[data$education == "5"] = "associates degree"
data$education[data$education == "6"] = "college degree"
data$education[data$education == "7"] = "master's degree"
data$education[data$education == "8"] = "doctoral degree"

#recode work status
data$`work status` = as.character(data$`work status`)
data$`work status`[data$`work status` == "1"] = "full time"
data$`work status`[data$`work status` == "2"] = "part time"
data$`work status`[data$`work status` == "3"] = "seasonally"
data$`work status`[data$`work status` == "4"] = "day labor"
data$`work status`[data$`work status` == "5"] = "unemployed"
```

## Demographical Variables

```{r age distribution}
#Plot age distribution
data %>%
  filter(Ethnics!="Other") %>%
  filter(!is.na(gender)) %>%
  ggplot() +
  geom_bar(aes(x=age,fill=gender),stat="count",position = "dodge")+
  theme_bw() +
  scale_fill_manual(values=c("deepskyblue1","deeppink1")) +
  labs(title = "Age Distribution",
       x = "Age",
       y = "Population")
#Calculate the mean and standard deviation of age
mean(data$age)
sd(data$age)
```

```{r Ethnics}
#Reorder gender and ethnics
data$gender <- factor(data$gender,levels = c("male", "female"))
data$Ethnics <- factor(data$Ethnics,levels = c("Caucasian", "African","Asian","Hispanic","Native","Pacific Island"))

#Plot the ethnics distribution
data %>%
  filter(Ethnics!="Other") %>%
  filter(!is.na(gender)) %>%
  ggplot() +
  geom_bar(aes(x=Ethnics,fill=gender),stat="count",position = "dodge")+
  theme_bw() +
  scale_fill_manual(values=c("deepskyblue1","deeppink1")) +
  labs(title="Distribution of Ethnicity by Gender",
       x="Ethnicity",
       y="Count")
```

```{r Service}
#reoeder gender, service type, and service branch
data$gender <- factor(data$gender,levels = c("male", "female"))
data$ServiceType <- factor(data$ServiceType,levels = c("Active Duty","Reserve","Guard"))
data$ServiceBranch <- factor(data$ServiceBranch,levels = c("Army","Air Force","Coast Guard","Marines","National","Navy"))

#plot the service type and service branch distribution
data %>%
  filter(!is.na(gender)) %>%
  ggplot() +
  geom_bar(aes(x=ServiceType,,fill=gender),stat="count",position = "dodge")+
  geom_bar(aes(x=ServiceBranch,fill=gender),stat="count",position = "dodge")+
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1)) +
  geom_vline(aes(xintercept=3.5),linetype="dashed") +
  scale_fill_manual(values=c("deepskyblue1","deeppink1")) +
  scale_x_discrete(limits=c("Active Duty",
                            "Reserve",
                            "Guard",
                            "Navy",
                            "Marines",
                            "Army",
                            "Air Force",
                            "Coast Guard",
                            "National"),drop=FALSE,na.translate=FALSE) +
  labs(title="Distribution of Service Type and Service Branch",
       x="Service",
       y="Count") +
  geom_text(aes(label="Service Type",x=2,y=3400)) +
  geom_text(aes(label="Service Branch",x=6.8,y=3400))
```

```{r SES}
#plot annual income distribution 
data %>%
  filter(!is.na(gender)) %>%
  ggplot() +
  geom_bar(aes(x=income,fill=gender),stat="count",position = "dodge") + 
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1)) +
  scale_fill_manual(values=c("deepskyblue1","deeppink1")) +
  scale_x_discrete(limits=c("less than 15k",
                            "15k~30k",
                            "30k~45k",
                            "45k~60k",
                            "60k~75k",
                            "75k~100k",
                            "more than 100k")) +
  labs(title="Distribution of Annual Income",
       subtitle="Unit: US Dollars",
       x="Annual Income",
       y="Count")

#plot education level distribution 
data %>%
  filter(!is.na(gender)) %>%
  ggplot() +
  geom_bar(aes(x=education,fill=gender),stat="count",position = "dodge") + 
  theme_bw() +
  theme(axis.text.x=element_text(angle=20, hjust=1)) +
  scale_fill_manual(values=c("deepskyblue1","deeppink1")) +
  scale_x_discrete(limits=c("some high school",
                            "GED",
                            "high school",
                            "some college",
                            "associates degree",
                            "college degree",
                            "master's degree",
                            "doctoral degree")) +
  labs(title="Distribution of Education Level",
       x="Education Level",
       y="Count")

#Plot work status and income distribution 
data %>%
  filter(!is.na(gender)) %>%
  ggplot() +
  geom_bar(aes(x=income,fill=gender),stat="count",position = "dodge") +
  geom_bar(aes(x=`work status`,fill=gender),stat="count",position = "dodge") + 
  theme_bw() +
  theme(axis.text.x=element_text(angle=20, hjust=1)) +
  geom_vline(aes(xintercept=7.5),linetype="dashed") +
  scale_fill_manual(values=c("deepskyblue1","deeppink1")) +
  scale_x_discrete(limits=c("less than 15k",
                            "15k~30k",
                            "30k~45k",
                            "45k~60k",
                            "60k~75k",
                            "75k~100k",
                            "more than 100k",
                            "unemployed",
                            "full time",
                            "part time",
                            "seasonally",
                            "day labor")) +
  labs(title="Distribution of Socioeconomic Status",
       x="Socioeconomic Status",
       y="Count") + 
  geom_text(aes(label="Income",x=3.4,y=640)) +
  geom_text(aes(label="Work Status",x=10.1,y=640))

#plot relationship status distribution 
data %>%
  filter(!is.na(gender)) %>%
  ggplot() +
  geom_bar(aes(x=relationship,fill=gender),stat="count",position = "dodge") + 
  theme_bw() +
  theme(axis.text.x=element_text(angle=20, hjust=1)) +
  scale_fill_manual(values=c("deepskyblue1","deeppink1")) +
  scale_x_discrete(limits=c("married",
                            "single",
                            "divorced",
                            "separated",
                            "cohabitating",
                            "remarried",
                            "widowed",
                            "civil union")) +
  labs(title="Distribution of Relationship Status",
       x="Relationship Status",
       y="Count")
```

```{r sociodemographic variable caculation}
#Create a table for Ethnics frequency and percentage
Ethnics.table <- data %>% 
  filter(!is.na(gender)) %>%
  filter(!is.na(Ethnics)) %>%
  count(Ethnics) %>%
  mutate(percent=n/sum(n))

#Create a table for education level and percentage
education.table <- data %>% 
  filter(!is.na(gender)) %>%
  filter(!is.na(education)) %>%
  count(education) %>%
  mutate(percent=n/sum(n))

#Create a table for service type and percentage
servicetype.table <- data %>% 
  filter(!is.na(gender)) %>%
  filter(!is.na(ServiceType)) %>%
  count(ServiceType) %>%
  mutate(percent=n/sum(n))

#Create a table for service branch and percentage
servicebranch.table <- data %>% 
  filter(!is.na(gender)) %>%
  filter(!is.na(ServiceBranch)) %>%
  count(ServiceBranch) %>%
  mutate(percent=n/sum(n))

#Create a table for income and percentage
income.table <- data %>% 
  filter(!is.na(gender)) %>%
  filter(!is.na(income)) %>%
  count(income) %>%
  mutate(percent=n/sum(n))

#Create a table for work status and percentage
workstatus.table <- data %>% 
  filter(!is.na(gender)) %>%
  filter(!is.na(`work status`)) %>%
  count(`work status`) %>%
  mutate(percent=n/sum(n))
```

## Analyses

### Correlation
```{r correlation}
#Load the package for heatmap
library(corrplot)
correlation <- data %>%
  select(`emotion support`,`depression`,`sleep quality`,`anxiety`,`alcohol`) %>%
  filter(!is.na(depression),
         !is.na(alcohol),
         !is.na(`sleep quality`),
         !is.na(anxiety),
         !is.na(`emotion support`))

#Correlation efficients matrix
cor.matrix <- round(cor(correlation),2)

#Get upper triangle of the correlation matrix
get_upper_tri <- function(cor.matrix){
  cor.matrix[lower.tri(cor.matrix)]<- NA
    return(cor.matrix)
}

upper_tri <- get_upper_tri(cor.matrix)

#Melt the correlation matrix
library(reshape2)
melted_cormat <- melt(upper_tri, na.rm = TRUE)
#Create a Heatmap
library(ggplot2)
ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low="deepskyblue",
                      high="coral",
                      mid="white",
                      midpoint=0, 
                      limit=c(-1,1), 
                      space="Lab", 
                      name="Pearson Correlation") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1,hjust = 1))+
  coord_fixed()+
  labs(x = "",
       y = "")+
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.8),
    legend.direction = "horizontal")+
  guides(fill=guide_colorbar(barwidth=7,barheight=1,title.position="top",title.hjust=0.5))

#Create a correlation plot with corrplot package
library(corrplot)  
cor.mtest(correlation)
```

### Emotional Support
```{r sleep quality~relationship+*gender}
#Establish the model for relationship status, gender, and sleep quality
model_rel_sleep=lm(`sleep quality`~relationship*gender,data)
anova(model_rel_sleep)

#Plot a graphic for relationship status, gender, and sleep quality
data %>%
  ##Choose major categories of relationship
  filter(relationship %in% c("divorced","single","married")) %>%
  ggplot()+
  geom_boxplot(aes(x=relationship,y = `sleep quality`,color = gender)) +
  labs(title = "Relationship Status and Sleep",
       x = "Relationship Status",
       y = "Sleep Quality") +
  ##Choose corresponding colors
  scale_color_manual(values=c("deepskyblue1","deeppink1")) +
  theme_bw()

#Create subsets for pairwise t-test
d.married <- data.relationship %>%
  filter(relationship=="married")
d.single <- data.relationship %>%
  filter(relationship=="single")
d.divorced <- data.relationship %>%
  filter(relationship=="divorced")

#Post-hoc pairwise t-test for sleep quality
pairwise.t.test(data.relationship$`sleep quality`,data.relationship$relationship,p.adj="none")
t.test(d.divorced$`sleep quality`,d.single$`sleep quality`)
t.test(d.divorced$`sleep quality`,d.married$`sleep quality`)
t.test(d.married$`sleep quality`,d.single$`sleep quality`)
```

```{r anxiety~relationship*gender}
#Establish the model for relationship status, gender, and anxiety
model_rel_anx=lm(data=data,formula=anxiety~gender*relationship)
anova(model_rel_anx)

#Plot a graphic for relationship status, gender, and anxiety
data%>%
  filter(relationship %in% c("divorced","single","married")) %>%
  ggplot() + 
  geom_boxplot(aes(x=relationship,y = anxiety, color = gender)) +
  labs(title = "Relationship Status and Anxiety",
       x = "Relationship Status",
       y = "Anxiety Level") +
  scale_color_manual(values=c("deepskyblue1","deeppink1")) +
  theme_bw()

#Post-hoc pairwise t-test for anxiety
t.test(d.divorced$anxiety,d.single$anxiety)
t.test(d.divorced$anxiety,d.married$anxiety)
t.test(d.married$anxiety,d.single$anxiety)
```

```{r depression~relationship*gender}
#Establish the model for relationship status, gender, and depression
model_rel_dep=lm(data=data,formula=depression~relationship*gender)
anova(model_rel_dep)

#Plot a graphic for relationship status, gender, and depression
data %>%  
  ggplot() + 
  geom_boxplot(aes(x=relationship,y = depression, color = gender)) +
  labs(title = "Relationship between Relationship Status and Depression",
       x = "Relationship Status",
       y = "Depression") +
  scale_color_manual(values=c("deepskyblue1","deeppink1")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))

#Post-hoc pairwise t-test for depression
t.test(d.divorced$depression,d.single$depression)
t.test(d.divorced$depression,d.married$depression)
t.test(d.married$depression,d.single$depression)
```

```{r alcohol usage~relationship*gender}
#Establish the model for relationship status, gender, and alcohol usage
model_rel_alc=lm(data=data,formula=alcohol~relationship*gender)
anova(model_rel_alc)

#Plot a graphic for relationship status, gender, and alcohol usage
data %>%  
  filter(relationship %in% c("divorced","single","married")) %>%
  ggplot() + 
  geom_boxplot(aes(x=relationship,y = alcohol, color = gender)) +
  labs(title = "Relationship Status and Alcohol Abuse",
       x = "Relationship Status",
       y = "Alcohol Usage") +
  theme_bw() +
  scale_color_manual(values=c("deepskyblue1","deeppink1"))

#Post-hoc pairwise t-test for alcohol usage
t.test(d.divorced$alcohol,d.single$alcohol)
t.test(d.divorced$alcohol,d.married$alcohol)
t.test(d.married$alcohol,d.single$alcohol)
```

```{r sleep quality~emotional support}
#Establish the model for sleep quality and emotional support
model_es_gender_sleep=lm(data=data,formula=`sleep quality`~ `emotion support`+gender)
anova(model_es_gender_sleep)

#Plot a graphic for sleep quality and emotional support
data %>%
  ggplot(aes(x=`emotion support`,y = `sleep quality`, color = gender))+
  geom_jitter(alpha=0.4, size=0.5)+
  geom_smooth(alpha=.15)+
  theme_bw()+
  scale_color_manual(values=c("deepskyblue1","deeppink1")) +
   labs(title = "Relationship between Emotional Support and Sleep Disorder",
       x = "Emotional Support (PROMIS score)",
       y = "Sleep Disorder (Insomnia Severity Index score)")
```

```{r anxiety~emotional support}
#Establish the model for anxiety and emotional support
model_es_gender_anx=lm(data=data,formula=anxiety ~ `emotion support`)
anova(model_es_gender_anx)

#Plot a graphic for anxiety and emotional support
data %>%
  ggplot(aes(x=`emotion support`,y=anxiety))+
  geom_jitter(alpha=0.4, size=0.5,color="deepskyblue4")+
  geom_smooth(alpha=.15)+
  theme_bw()+
  labs(title = "Relationship between Emotional Support and Anxiety",
       x = "Emotional Support (PROMIS score)",
       y = "Anxiety (GAD-7 Score)")
```

```{r alcohol~emotional support}
#Establish the model for alcohol usage and emotional support
model_es_gender_anx=lm(data=data,formula=alcohol ~ `emotion support`)
anova(model_es_gender_anx)

#Plot a graphic for alcohol usage and emotional support
data %>%
  ggplot(aes(x=`emotion support`,y=alcohol))+
  geom_jitter(alpha=0.4, size=0.5,color="deepskyblue4")+
  geom_smooth()+
  theme_bw()+
  labs(title = "Relationship between Emotional Support and Alcohol Use",
       x = "Emotional Support (PROMIS score)",
       y = "Alcohol Use (AUDIT score)")
```

```{r depression~emotional support}
#Establish the model for depression and emotional support
model_es_gender_depression=lm(data=data,formula=depression ~ `emotion support`+gender)
anova(model_es_gender_depression)

#Plot a graphic for depression and emotional support
data %>%
  ggplot(aes(x=`emotion support`,y=depression))+
  geom_jitter(alpha=0.4, size=0.5,color="deepskyblue4")+
  geom_smooth()+
  theme_bw()+
  labs(title = "Relationship between Emotional Support and Depression",
       x = "Emotional Support (PROMIS score)",
       y = "Depression (PHQ-9 score)")
```

### Socioeconomic Status and Mental Health
```{r anxiety ~ income}
#Establish the model for anxiety and income
model_inc_anx= data %>% filter(!is.na(income)) %>% lm(.,formula=anxiety ~ income+gender)
anova(model_inc_anx)
data %>% group_by(income) %>% summarise(n = n())

#Plot a graphic for anxiety and income
data %>%
  filter(!is.na(income)) %>% 
  ggplot() + 
  geom_boxplot(aes(x= income,y = anxiety, color = gender)) +
  labs(title = "Relationship between Income Level and Anxiety",
       x = "Income level",
       y = "anxiety") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))
```

```{r anxiety ~ education}
#Establish the model for anxiety and education
model_edu_anx= data %>% filter(!is.na(education)) %>% lm(.,formula=anxiety ~ education+gender) 
anova(model_edu_anx)
summary(model_edu_anx)
data %>% group_by(income) %>% summarise(n = n())

#Plot a graphic for anxiety and education
data %>%
  filter(!is.na(education)) %>% 
  ggplot() + 
  geom_boxplot(aes(x= education,y = anxiety, color = gender)) +
  labs(title = "Relationship between Education Level and Anxiety",
       x = "Education level",
       y = "anxiety") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))
```

```{r alcohol ~ income}
#Establish the model for alcohol and income
model_edu_alc= data %>% filter(!is.na(income)) %>% lm(.,formula=alcohol ~ income + gender) 
anova(model_edu_alc)
summary(model_edu_alc)
data %>% group_by(income) %>% summarise(n = n())

#Plot a graphic for alcohol and income
data %>%
  filter(!is.na(alcohol)) %>% 
  filter(!is.na(income)) %>%
  ggplot() + 
  geom_boxplot(aes(x= income,y = alcohol, color = gender)) +
  labs(title = "Relationship between Income Level and Alcohol Usage",
       x = "Income level",
       y = "alcohol") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))
```

```{r anxiety ~ work status}
#Establish the model for anxiety and work status
model_work_anx= data %>% filter(!is.na(`work status`)) %>% lm(.,formula=anxiety ~ `work status` + gender) 
anova(model_work_anx)
summary(model_work_anx)
data %>% group_by(`work status`) %>% summarise(n = n())

#Plot a graphic for anxiety and work status
data %>%
  filter(!is.na(`work status`)) %>% 
  ggplot() + 
  geom_boxplot(aes(x= `work status`,y = anxiety, color = gender)) +
  labs(title = "Relationship between Work Status and Anxiety",
       x = "work status",
       y = "anxiety") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=30, hjust=1))
```

### ANOVAs for SES and mental health
```{r ANOVAs for SES and mental health}
#Establish the model for anxiety and income
model_inc_anx= data %>% filter(!is.na(income)) %>% lm(.,formula=anxiety ~ income)
anova(model_inc_anx)

#Establish the model for sleep quality and income
model_inc_sleep= data %>% filter(!is.na(income)) %>% lm(.,formula=`sleep quality` ~ income)
anova(model_inc_sleep)

##Establish the model for depression and income
model_inc_dep= data %>% filter(!is.na(income)) %>% lm(.,formula=depression ~ income)
anova(model_inc_dep)

#Establish the model for alcohol usage and income
model_inc_alc= data %>% filter(!is.na(income)) %>% lm(.,formula=alcohol ~ income)
anova(model_inc_alc)

#Establish the education frequency table
data %>% group_by(education) %>% count(education)
#Establish the model for anxiety and education
model_edu_anx= data %>% ilter(!is.na(education)) %>% lm(.,formula=anxiety ~ education)
anova(model_edu_anx)

#Establish the model for sleep quality and education
model_edu_sleep= data %>% filter(!is.na(education)) %>% lm(.,formula=`sleep quality` ~ education)
anova(model_edu_sleep)

#Establish the model for depression and education
model_edu_dep= data %>% filter(!is.na(education)) %>% lm(.,formula=depression ~ education)
anova(model_edu_dep)

#Establish the model for alcohol usage and education
model_edu_alc= data %>% filter(!is.na(education)) %>% lm(.,formula=alcohol ~ education)
anova(model_edu_alc)

#Establish the education frequency table
data %>% group_by(`work status`) %>% count(`work status`)
#Establish the model for anxiety and work status
model_work_anx= data %>% filter(!is.na(`work status`)) %>% lm(.,formula=anxiety ~ `work status`)
anova(model_work_anx)

#Establish the model for sleep quality and work status
model_work_sleep= data %>% filter(!is.na(`work status`)) %>% lm(.,formula=`sleep quality` ~ `work status`)
anova(model_work_sleep)

#Establish the model for depression and work status
model_work_dep= data %>% filter(!is.na(`work status`)) %>% lm(.,formula=depression ~ `work status`)
anova(model_work_dep)

#Establish the model for alcohol usage and work status
model_work_alc= data %>% filter(!is.na(`work status`)) %>% lm(.,formula=alcohol ~ `work status`)
anova(model_work_alc)
```