---
title: "Veteran's Life History and Psychological Health"
subtitle: "A dataset from San Diego VA hospital"
author: "Shuai Shao, Shuyuan Shi, Xinyue Li"
date: "11/14/2019"
output: html_document
---
##Data management
In this group project, we will use github to record and track every team member's work. You can find our project at [https://github.com/shaoshuai95/PSYC201A-project.git].

We have get the permission to access to the data from VA hospital. Currently, we have... varables

We will gather more variables such as veteran's social support, traumatical experience... (such as childhood trauma), 

We plan to analyze:
1) hypothesis + how to test it (method: linear regression.../chi-squred test...)
2)
...

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
```

```{r import data}
data <- readxl::read_xlsx("PSYC201a.xlsx")
```

```{r clean the data}
# select varables of interest
data <- data %>% 
  select(-2,-13)

# rename the columns
data <- rename(data,"gender"="demo_gender",
                    "Hispanic"="demo_ethnic",
                    "Caucasian"="demo_racewhite",
                    "African"="demo_race_black",
                    "Native"="demo_race_amind",
                    "Pacific Island"="demo_race_pacisl",
                    "Asian"="demo_race_asian",
                    "Other"="demo_race_oth",
                    "Unclassified"="demo_race_decline",
                    "Age"="Calculate999age",
                    "Relationship"="demo_relationship",
                    "ServiceType"="serv_type_0",
                    "ServiceBranch"="serv_branch_0",
                    "ServiceDeployCount"="serv_deploy_count") 
```

```{r dealing with missing data}
# replace all "999"s with "NA"
data[data==999]=NA
```

```{r filter out unreasonable cases}
# filter out unreasonable cases
data <- data %>% 
  filter(Age<100) %>%
  filter(Age>18)
```

```{r recode gender }
# male=1, female=0
data$gender[data$gender==2]=0
```

```{r rename values}
#rename relationships
data$Relationship=as.character(data$Relationship)
data$Relationship[data$Relationship=="1"]="single"
data$Relationship[data$Relationship=="2"]="married"
data$Relationship[data$Relationship=="3"]="seperated"
data$Relationship[data$Relationship=="4"]="divorced"
data$Relationship[data$Relationship=="5"]="cohabitating"
data$Relationship[data$Relationship=="6"]="civil union"
data$Relationship[data$Relationship=="7"]="remarried"
data$Relationship[data$Relationship=="8"]="widowed"

#recode service type
data$ServiceType = as.character(data$ServiceType)
data$ServiceType[data$ServiceType == "1"] = "Active Duty"
data$ServiceType[data$ServiceType == "2"] = "Reserve"
data$ServiceType[data$ServiceType == "3"] = "Guard"

#recode service branch
data$ServiceBranch = as.character(data$ServiceBranch)
data$ServiceBranch[data$ServiceBranch == "1"] = "Army"
data$ServiceBranch[data$ServiceBranch == "2"] = "Air Force"
data$ServiceBranch[data$ServiceBranch == "3"] = "Coast Guard"
data$ServiceBranch[data$ServiceBranch == "4"] = "Marines"
data$ServiceBranch[data$ServiceBranch == "5"] = "National"
data$ServiceBranch[data$ServiceBranch == "6"] = "Navy"
```

```{r demographical variables}
#gender and ethnics distribution

#ethnic distribution

#age ditribution

#gender and ethnic

```

